### 2020-02-16 paper.tex
fb 1st pass complete

\documentclass[runningheads]{llncs}
\usepackage{times}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
%\usepackage[ruled,linesnumbered,noend,oldcommands]{algorithm2e}
\usepackage{xspace}
\usepackage{mathtools}
\usepackage{graphicx}
\usepackage{array}
\usepackage{xcolor}
\usepackage{theorem}
\usepackage[T1]{fontenc}
% \usepackage[T1,hyphens]{url}
\usepackage{hyperref}
%\PassOptionsToPackage{hypnens}{url}
\title{Clause size reduction with all-UIP Learning}
\author{Nick Feng \and Fahiem Bacchus}
\authorrunning{N. Feng and F. Bacchus}
\institute{Department of Computer Science, University of Toronto, Canada, 
  \email{\{nfeng,fbacchus\}@cs.toronto.edu}}

\algrenewcommand\algorithmicprocedure{}
\algrenewcommand\algorithmicthen{}
\algrenewcommand\algorithmicdo{}

\newcommand{\fs}[1]{\fontsize{#1}{#1}\selectfont}
\newcommand{\fss}[2]{\fontsize{#1}{#2}\selectfont}
\newcommand{\sat}{SAT\xspace}
\newcommand{\code}[1]{\text{#1}}
\newcommand{\assertionTrail}{\trail}
\newcommand{\LBD}{\text{LBD}\xspace}
\newcommand{\gap}{\text{Gap}}
\newcommand{\tgap}{t_{\mathit{gap}}}
\newcommand{\ntries}{n_{\mathit{tries}}}
\newcommand{\nsuc}{n_{\mathit{succ}}}
\newcommand{\allUip}{\textit{stable-allUIP}}
\newcommand{\tryuiplevel}{\textit{try-uip-level}\xspace}
\newcommand{\allUipPure}{\textit{pure-allUIP}\xspace}
\newcommand{\allUipMin}{\textit{min-allUIP}\xspace}
\newcommand{\allUipAct}{\textit{allUIP-Active}}
\newcommand{\allUipIn}{\textit{allUIP-Inclusive}}
\newcommand{\allUipEx}{\textit{allUIP-Exclusive}}
\newcommand{\MapleBase}{\textit{MapleCOMSPS\_LRB}}
\newcommand{\MapleSeven}{\textit{MapleLCMDist}}
\newcommand{\MapleNine}{\textit{MapleLCMDiscChronoBT-DL-v3} }
\newcommand{\MapleEight}{\textit{MapleLCMDiscChronoBT}}
\newcommand{\MapleNineShort}{\textit{MapleCB-DL} }
\newcommand{\expSAT}{\textit{expMaple\_CM\_GCBumpOnlyLRB} }
\newcommand{\expSATShort}{\textit{expMaple} }
\newcommand{\MapleIUIPPure}{\text{Maple-\allUipPure}}
\newcommand{\MapleIUIMin}{\text{Maple-\allUipMin}}
\newcommand{\MapleEightShort}{\textit{MapleCB}}
\newcommand{\Set}[2]{\{\,#1\mid#2\,\}}
\newcommand{\Array}[2]{[\,#1\mid#2\,]}
\newcommand{\target}{\textbf{Succ_t}}
\newcommand{\trail}{\ensuremath{\mathcal{T}}}
\newcommand{\trailIdx}[1]{\ensuremath{\iota(#1)}}
\newcommand{\dlevel}[1]{\ensuremath{\mathit{decLvl}(#1)}}
\newcommand{\dlevels}{\ensuremath{\mathit{decLvls}}}
\newcommand{\var}{\text{var}}
\newcommand{\true}{\textsc{true}\xspace}
\newcommand{\false}{\textsc{false}\xspace}
\newcommand{\reason}[1]{\ensuremath{\mathit{reason}(#1)}}
\newcommand{\resolve}{\bowtie}
\renewcommand{\implies}{\rightarrow}
\newcommand{\ctry}{C_{\mathit{try}}}
\newcommand{\type}{\textit{type}}
\newcommand{\unmarked}{\textbf{unmarked}}
\newcommand{\mrk}{\textbf{mark}}
\setlength{\theorempreskipamount}{3pt}
\setlength{\theorempostskipamount}{3pt}

\newtheorem{Obs}{Observation}
\newtheorem{Cor}{Corollary}
\newtheorem{defn}{Definition}
\newtheorem{thm}{Theorem}

\newcommand{\whitebox}{\raisebox{.5ex}{\fbox{\hspace*{.2ex}}}}

%Comments
\newcommand{\nf}[1]{{\color{red}{#1}}}
\newcommand{\fb}[1]{{\color{blue}{#1}}}

\begin{document}
\maketitle              % typeset the header of the contribution
% 
\begin{abstract}
\end{abstract}

\section{Introduction}
Clause learning is an essential technique in \sat solvers. There is good
evidence to indicate that it is, in fact, the most important technique
used in modern \sat solvers \cite{DBLP:conf/sat/KatebiSS11}. In early
\sat research a number of different clause learning techniques were
proposed
\cite{DBLP:conf/iccad/ZhangMMM01,DBLP:conf/iccad/SilvaS96,DBLP:journals/tc/Marques-SilvaS99,DBLP:conf/aaai/BayardoS97}.
However, following the revolutionary performance improvements achieved
by the Chaff \sat solver, the field has converged on using the 1-UIP
(first Unique Implication Point) scheme
\cite{DBLP:conf/iccad/ZhangMMM01} employed in Chaff
\cite{DBLP:conf/dac/MoskewiczMZZM01} (as well as other techniques
pioneered in the Chaff solver).\footnote{The idea of UIP clauses was
  first mentioned in \cite{DBLP:journals/tc/Marques-SilvaS99}, and
  1-UIP clauses along with other UIP clauses were learnt and used in
  the earlier GRASP \sat solver.} Since then almost all \sat solvers
have employed the 1-UIP clause learning scheme, along with clause
minimization \cite{DBLP:conf/sat/SorenssonB09}, as their primary
method for learning new clauses.

However, other clause learning schemes can be used in SAT solvers
without changes to their main data structures. Furthermore, advances
in our understanding allow us to better understand the potential
advantages and disadvantages of these alternate schemes. In this paper
we reexamine some of these previous clause learning schemes, with a
focus on the schemes described in
\cite{DBLP:conf/iccad/ZhangMMM01}. Improved understanding of \sat
solvers, obtained from the last decade of research, allows us to see
that in their original form these other clause learning schemes suffer
significant disadvantages over 1-UIP clause learning.

One of the previously proposed schemes was the all-UIP scheme
\cite{DBLP:conf/iccad/ZhangMMM01}. In this paper we propose a new way
to exploit the main ideas of this scheme that avoids its main
disadvantages. In particular, we propose to use a all-UIP like clause
learning scheme to generate smaller learnt clauses which retain the
good properties of standard 1-UIP clauses. Our method is related to,
but not the same as, various clause mininimization methods that try to
remove redundant literals from the 1-UIP clause yielding a clause that
is a subset of the 1-UIP clause, e.g.,
\cite{DBLP:conf/sat/SorenssonB09,DBLP:conf/ijcai/LuoLXML17,DBLP:conf/sat/WieringaH13}.
Our method is orthogonal to clause minimization. In particular, our
approach can learn a clause that is not a subset of 1-UIP clause but
which still serves all of the same purposes as the 1-UIP clause.
Various minimization techniques can be applied on top of our method to
remove redundant literals from the clauses we learn.

We present various versions of our method and show that these variants
are often capable of learning shorter clauses that the 1-UIP scheme,
and that this can lead to useful performance gains in state of the art
\sat solvers.

\section{Clause Learning Framework}
We first provide some background and a framework for understanding
clause learning as typically used in CDCL \sat solvers. A
propositional formula $F$ expressed in Conjunctive Normal Form (CNF),
contains a set of variables $V$. A literal is a variable $v\in V$ or
its negation $\lnot v$. For a literal $\ell$ we let $\var(\ell)$
denote its underlying variable. A CNF consists of a conjunction of
clauses, each of which is a disjunction of literals. We often view a
clause as being a set of literals and employ set notation, e.g.,
$\ell\in C$ and $C'\subset C$.

Two clauses $C_1$ and $C_2$ can be \emph{resolved} when they contain
\emph{conflicting} literals $\ell\in C_1$ and $\lnot \ell \in
C_2$. Their resolvant $C_1 \resolve C_2$ is the new clause
$(C_1 \cup C_2) - \{\ell, \lnot \ell\}$. The resolvant will be a
tautology (i.e., a clause containing a literal $l$ and its negation
$\lnot l$) if $C_1$ and $C_2$ contain more than one pair of
conflicting literals.

We assume the reader is familiar with the operations of CDCL \sat
solvers, and the main data structures used in such solvers. A good
source for this background is \cite{DBLP:series/faia/SilvaLM09}.

\paragraph{The Trail.}
CDCL \sat solvers maintain a \textbf{trail}, $\trail$, which is a
\textit{non-contradictory, non-redundant sequence of literals} that
have been assigned \true by the solver; i.e.
$\ell\in\trail \implies \lnot\ell \not\in\trail$, and $\trail$
contains no duplicates. Newly assigned literals are added to the end
of the trail, and on backtrack literals are removed from the end of
the trail and unassigned.  If literal $\ell$ is on the trail let
$\trailIdx{\ell}$ denote its index on the trail, i.e,
$\trail[\trailIdx{\ell}] = \ell$. For convenience, we also let
$\trailIdx{\ell} = \trailIdx{\lnot \ell} = \trailIdx{\var(\ell)}$ even
though neither $\lnot \ell$ nor $\var(\ell)$ are actually on
$\trail$. If $x$ and $y$ are both on the trail and
$\trailIdx{x} < \trailIdx{y}$ we say that \textit{$x$ appears before
  $y$ on the trail}.

Two types of true literals appear on the trail: \emph{decision
  literals} that have been assumed to be true by the solver, and
\emph{unit propagated literals} that are forced to be true because
they are the sole remaining unfalsified literal of a clause. Each
literal $\ell\in\trail$ has a decision level $\dlevel{\ell}$ which is
equal to the number of decision literals appearing before $\ell$ on
the trail plus one; hence, $\dlevel{d}=1$ for the first decision
literal $d\in\trail$.  The set of literals on $\trail$ that have the
same decision level forms a contiguous subsequence of $\trail$ that
starts with a decision literal $d_i$ and ends just before the next
decision literal $d_{i+1}$. If $\dlevel{d_i} = i$ we call this
subsequence of $\trail$ the \textit{$i$-th decision level}.

Each literal $\ell\in\trail$ also has a clausal reason
$\reason{\ell}$. If $\ell$ is a unit propgated literal,
$\reason{\ell}$ is a clause of the formula such that
$\ell \in \reason{\ell}$ and
$\forall x \in \reason{\ell}.\, x\neq \ell \implies (\lnot x \in
\trail \land \trailIdx{\lnot x} < \trailIdx{\ell}$. That is,
$\reason{\ell}$ is a clause that has become unit implying $\ell$ due
to the literals on the trail above $\ell$. If $\ell$ is a decision
literal then $\reason{\ell} = \varnothing$.

In most \sat solvers clause learning is initiated as soon as a clause
is falsified by $\trail$. In this paper we will be concerned with the
subsequent clause learning process which uses $\trail$ to derive a new
clause. In some implementations the trail might be altered during
clause learning. Here, however, we will assume that $\trail$
\emph{remains intact during clause learning} and is only changed after
the new clause is learnt. After the new clause is learnt the $\trail$
will be changed by backtracking.

Say that $\trail$ falsifies a clause $C_I$, and that the last decision
literal $d_k$ in $\trail$ has decision level $k$. Consider
$\trail_{k-1}$ the prefix of $\trail$ above the last decision level,
i.e., the sequence of literals
$\trail[0]$---$\trail[\trailIdx{d_k}-1]$. We will assume that
$\trail_{k-1}$ is \textbf{propagation complete}, although the full
trail $\trail$ might not be. This means that (a) no clause was
falsified by $\trail_{k-1}$. And (b) if $C_u$ is a clause containing
the literal $x$ and all literals in $C_u$ except for $x$ are falsified
by $\trail_{k-1}$, then $x\in \trail_{k-1}$ and
$\dlevel{x} \leq\max\{\dlevel{y} | y\in C_u \land y\neq x\}$. This
means that if $x$ appears in a clause made unit it must have been
added to the trail, and added at or before decision level the clause
became unit. Note that more than one clause might be made unit by
$\trail$ forcing $x$, or $x$ might be set as a decision before being
forced. This condition ensures that $x$ appears in $\trail$ at or
before the first decision level it is forced by any clause.

Any clause falsified by $\trail$ is called a \textbf{conflict}. When a
conflict is found, the final level of the trail, $k$, need not be
propagation complete as the solver typically stops propagation as soon
as it finds a conflict. This means that (a) other clauses might be
falsified by $\trail$ besides the conflict found, and (b) other
literals might be unit implied by $\trail$ but not added to $\trail$.

\begin{defn}[Trail Resolvant]
    A trail resolvant is a clause arising from resolving a conflict
    against the reason clause of some literal $\ell \in \trail$. Every
    trail resolvant is also a conflict.
\end{defn}

The following things can be noted about trail resolvants: (1) trail
resolvants are never tautological, as the polarity of all literals in
$\reason{\ell}$ other than $\ell$ must agree with the polarity of all
literals in the conflict (they are all falsified by $\trail$); (2) one
polarity of the variable $\var(\ell)$ resolved on must be a unit
propagated literal whose negation appears in the conflict; and (3) any
variable in the conflict that is unit progagated in $\trail$ can be
resolved upon (the variable must appear in different polarities in the
conflict and in $\trail$).

\begin{defn}[Trail Resolution]
    A trail resolution is a sequence of trail resolvants applied to an
    initial conflict $C_I$ yielding a new conflict $C_L$.  A trail
    resolution is \textbf{ordered} if the sequence of variables $v_1$,
    \dots, $v_m$ resolved have strictly decreasing trail indicies:
    $\trailIdx{v_{i+1}} < \trailIdx{v_i}$ ($1\leq i < m$). (Note that
    this implies that no variable is resolved on more than once).
\end{defn}

Ordered trail resolutions resolve unit propagated literals from the
end of the trail to the beginning. Without lost of generality, we can
require that all trail resolutions be ordered.

\begin{Obs}
    If the unordered trail resolution $U$ yields the conflict clause
    $C_L$ from an initial conflict $C_I$, then there exists an ordered
    trail resolution $O$ that yields a conflict clause $C'_L$ such
    that $C'_L\subseteq C_L$.
\end{Obs}
\noindent
\textbf{Proof.} Let $U$ be the sequence of clauses $C_I = C_0$, $C_1$,
$\ldots$, $C_m= C_L$ obtained by resolving on the sequence of
variables $v_1$, $\ldots$, $v_m$ whose corresponding literals on
$\trail$ are $l_1$, $\ldots$, $l_m$. Reordering these resolution steps
so that the variables are resolved in order of decreasing trail index
and removing duplicates yields an ordered trail resolution $O$ with
the desired properties. Since no reason clause contains literals with
higher trail indicies, $O$ must be a valid trail resolution if $U$
was, and furthermore $O$ yields the clause
$C'_L = \bigcup_{i=1}^m \reason{l_i} - \{l_1, \lnot l_1, \ldots, l_m,
\lnot l_m\}$. Since $U$ resolves on the same variables (in a different
order) using the same reason clauses we must have $C'_L\subseteq
C_L$. It can, however, be the case that $C'_L$ is proper subset of
$C_L$: if $l_i$ is resolved away it might be reintroduced when
resolving on $l_{i+1}$ if $\trailIdx{l_{i+1}}  > \trailIdx{l_i}$.
\whitebox

The relevance of trail resolutions is that all proposed clause
learning schemes we are aware of use trail resolutions to produce
learnt clauses. Furthermore, the commonly used technique for clause
minimization \cite{DBLP:conf/sat/SorenssonB09} is also equivalent to a
trail resolution that yields the minimized clause from the
un-minimized clause. Interestingly, it is standard in \sat solver
implementations to perform resolution going backwards along the
trail. That is, these implementations are typically using ordered
trail resolutions. The above observation shows that this in fact the
right way to do this.

Ordered trail resolutions are a special case of \textit{trivial
  resolutions} \cite{DBLP:journals/jair/BeameKS04}. Trail resolutions
are specific to the trail data structure typically used in \sat
solvers. If $\trail$ falsifies a clause at its last decision level,
then its associated implication graph
\cite{DBLP:journals/tc/Marques-SilvaS99} contains a conflict
node. Cuts in the implication graph that separate the conflict from
the rest of the graph correspond to conflict clauses
\cite{DBLP:journals/jair/BeameKS04}. It is not difficult to see that
the proof Proposition~4 of \cite{DBLP:journals/jair/BeameKS04} applies
also to trail resolutions. This means that \textit{any conflict clause
  in the trail's implication graph can be derived using a trail
  resolution.}

\subsection{Some alternate Clause Learning Schemes}
A number of different clause learning schemes for generating a new
learnt clause from the initial conflict have been presented in prior
work, e.g.,
\cite{DBLP:conf/iccad/ZhangMMM01,DBLP:conf/iccad/SilvaS96,DBLP:journals/tc/Marques-SilvaS99,DBLP:conf/aaai/BayardoS97}.
Figure~\ref{fig:cl_schemes} gives a specification of some of these
methods: (a) the all-decision scheme which resolves away all implied
literals leaving a learnt clause over only decision literals; (c) the
1-UIP scheme which resolves away literals from the deepest decision
level leaving a learnt clause with a single literal at the deepest
level; (d) the all-UIP scheme which resolves away literals from each
decision level leaving a learnt clause with a single literal at each
decision level; and (e) the $i$-UIP scheme which resolves away
literals from the $i$ deepest decision levels leaving a learnt clause
with a single literal at its $i$ deepest decision levels. It should be
noted that when resolving away literals at decision level $i$ new
literals at decision levels less than $i$ might be introduced into the
clause. Hence, it is important in the $i$-UIP and all-UIP schemes to
use ordered trail resolutions.

Both the all-decision and all-UIP schemes yield a clause with only one
literal at each decision level, and the all-UIP clause will be no
larger that the all-decision clause. Furthermore, it is known
\cite{DBLP:journals/tc/Marques-SilvaS99} that once we reduce the
number of literals at a decision level $d$ to one, we could continue
performing resolutions and later achieve a different single literal at
the level $d$. In particular, a decision level might contain more than
one unique implication point. The algorithms given in
Figure~\ref{fig:cl_schemes} stop at the first UIP of a level, except
for the all-decision schemes with stops at the last UIP of each level.

\begin{figure}[t!]
{\fss{8pt}{10pt}
\begin{tabular}[t]{|l|c|l|}
  \cline{1-1}\cline{3-3}
  \begin{minipage}[t]{0.48\textwidth}
    \textbf{(a) All Decision Clause\rule{0pt}{1.1\topskip}}
    \begin{algorithmic}[0]
        \State \hspace*{-1em}\textbf{all\_decision}($C_I$)
        \State $C\gets C_I$
        \While{$\{l\,|\, l\in C \land\reason{l} \neq\varnothing\} \neq \emptyset$}
        \State \begin{tabular}{ll}
                 $\ell \gets \mbox{}$ & literal with highest trail index in\\
                                      & $\{l\,|\, l\in C \land\reason{l} \neq\varnothing\}$
               \end{tabular}
        \State $C \gets C\resolve \reason{\lnot \ell}$
        \EndWhile
        \State \textbf{return} $C$
    \end{algorithmic}
  \end{minipage}
  & &
  \begin{minipage}[t]{0.475\textwidth}
      \textbf{(b) Make level $i$ contain a single literal\rule{0pt}{1.1\topskip}}
      \begin{algorithmic}[0]
          \State \hspace*{-1em} \textbf{UIP\_level}($C$, $i$)
          \While{$\left|\left\{\ell \left| \begin{array}{l}\ell \in C\\
                                           \mbox{}\land \reason{l} \neq \varnothing \\
                                           \mbox{}\land \dlevel{l} = i\end{array}\right.\right\}\right| > 1$}
          \State \begin{tabular}{ll}
                   $l \gets \mbox{}$ & literal with highest trail index in\\
                                        & $\left\{\ell \left| \begin{array}{l}\ell \in C
                                          \land \reason{l} \neq \varnothing \\
                                          \mbox{}\land \dlevel{l} = i\end{array}\right.\right\}$
                 \end{tabular}
          \State $C \gets C\resolve \reason{\lnot l}$
          \EndWhile
          \State \textbf{return} $C$\strut
      \end{algorithmic}
  \end{minipage}
  \\\cline{1-1}\cline{3-3}
   \multicolumn{1}{c}{\vspace*{-7pt}}\\\cline{1-1}\cline{3-3}
  \begin{minipage}[t]{0.475\textwidth}
    \begin{minipage}{\textwidth}
        \textbf{(c) First UIP Clause \rule{0pt}{1.1\topskip}}
        \begin{algorithmic}[0]
            \State \hspace*{-1em}\textbf{1-UIP}($C_I$)
            \State $i\gets \max\{\dlevel{l}\, |\, l \in C_I\}$
            \State \textbf{return} \textbf{UIP\_level}($C_I$, i)
        \end{algorithmic}
        \vspace*{5pt}
    \end{minipage}\\[5pt]
    \hspace*{-1.8pt}\rule{1.03\textwidth}{.4pt}\\[5pt]
    \begin{minipage}{\textwidth}
        \textbf{(d) All UIP Clause \rule{0pt}{1.1\topskip}}
        \begin{algorithmic}[0]
          \State \hspace*{-1em}\textbf{all-UIP}($C_I$, $i$)
          \State $i = \bigl|\{\dlevel{l} \,|\, l \in \trail\}\bigl|$
          \State \Comment{$i$ is large enough to ensure all levels are UIP}
          \State \textbf{return} \textbf{$i$-UIP}(C, i)

      \end{algorithmic}
      \end{minipage}
  \end{minipage}
 & &
  \begin{minipage}[t]{0.48\textwidth}
      \vspace*{-20pt}
      \textbf{(e) $i$-UIP Clause}
      \begin{algorithmic}[0]
          \State \hspace*{-1em}\textbf{$i$-UIP}($C_I$, $i$)
          \State $C\gets C_I$
          \State $d\gets \max\{\dlevel{l}\,|\, l \in C\}$
          \For{($j \gets 1$; $j\leq i$; $j\gets j+1$)}
          \State \textbf{if} ($d = \varnothing$): \textbf{break}
          \State $C\gets\mbox{}$ \textbf{UIP\_level(C, d)}
          \State $d\gets \max\left\{\dlevel{l} \left|\begin{array}{l}l\in C \\ \mbox{} \land \dlevel{l}<d
                                                  \end{array}\right.\right\}$
          \EndFor
          \State \textbf{return} $C$\strut
      \end{algorithmic}
      \vspace*{3pt}
      \textit{Maximum of an empty set is $\varnothing$ \strut}
  \end{minipage}
\\\cline{1-1}\cline{3-3}
\end{tabular}
}
\caption{Some different clause learning schemes. All use the
      current trail $\trail$ and take as input an initial clause $C_I$
      falsified by $\trail$ at its deepest level.\label{fig:cl_schemes}}
\end{figure}

 %   \floatstyle{boxed}
 %   \restylefloat{figure}

\subsection{Asserting Clauses and LBD---Reasons to prefer 1-UIP
  clauses}
An \textbf{asserting clause} \cite{DBLP:journals/ai/PipatsrisawatD11}
is a conflict clause $C_L$ that has exactly one literal $\ell$ at its
deepest level, i.e.,
$\forall x\in C_L. \dlevel{x} \leq \dlevel{\ell} \land (\dlevel{x} =
\ell \implies x = \ell$). All of the clause learning schemes in
Figure~\ref{fig:cl_schemes} produced learnt clauses that are asserting.

The main advantage of asserting clauses is that they are 1-Empowering
\cite{DBLP:journals/ai/PipatsrisawatD11}, i.e., they allow unit
propagation to derive a new forced literal. Hence, asserting clauses
can be used to guide backtracking---the solver can backtrack from the
current deepest level to the point the learnt clause first becomes
unit and then use the learnt clause to add a new unit implicant to the
trail. Since all but the deepest level was propagation complete, this
means that the asserting clause must be a brand new clause; otherwise
that unit implication would already have been made. On the other hand,
if the learnt clause $C_L$ is not asserting then \emph{it could be
  that it is a duplicate of another clause already in the formula}. In
particular, $C_L$ contains at least two literals at the deepest level
If two of its literals at the deepest level were were not completely
unit propagated (unit propagation is aborted as soon as a conflict is
found), then $C_L$ could already be in the formula: it's two watches
might not be fully propagated and $C_L$ could be a falsified clause
not detected by the solver. (In general, more than one clause might be
falsified by $\trail$ and the \sat solver will generally stop when it
finds the first one).

The LBD of the learnt clause $C_L$ is the number of different decision
levels in it:
$\LBD(C_L)=\left|\big\{\dlevel{l}\,|\,l \in C_L\big\}\right|$
\cite{DBLP:conf/ijcai/AudemardS09}. Empirically LBD is a successful
predictor of clause usefulness: clauses with lower LBD tend to be more
useful. As noted in \cite{DBLP:conf/ijcai/AudemardS09}, from the
initial falsified clause $C_I$ the 1-UIP scheme will produce a clause
$C_L$ whose LBD is minimum among all asserting clauses that can be
learnt from $C_I$. If $C'$ is a trail resolvant of $C$ and a reason
clause $\reason{l}$, then $\LBD(C') \geq \LBD(C)$ since $\reason{l}$
must contain at least one other literal with the same decision level
as $l$ and might contain literals with decision levels not in
$C$. That is, each trail resolution step can only increase the \LBD of
the learnt clause. Hence, the 1-UIP scheme yields an asserting clause
with minimum LBD as it performs the minimum number of trail resolutions
required to generate an asserting clause.

The other schemes must perform more trail resolutions. In fact, all of
these schemes (all-decision, all-uip, i-uip) use trail resolutions in
which the 1-UIP clause appears. That is, they all must first generate
the 1-UIP clause and then continue with further trail resolution
steps. These extra resolution steps can introduce many addition
decision levels into the final clause. Hence, these schemes yield
learnt clauses whose LBD can be larger, and never smaller, that the
1-UIP clause.

Putting these two observations together we see that the 1-UIP scheme
produces asserting clauses with lowest possible LBD. This is a
compelling reasons for using this scheme. Hence, it is not surprising
that modern \sat sovlers almost exclusively use 1-UIP clause
learning.\footnote{Knuth in his sat13 CDCL solver \cite{Knuth:Sat13}
  uses an all-decision clause when the 1-UIP clause is too large. In
  this context an all-UIP clause could also be used as it would be no
  larger than the all decision clause.}

\section{Using all-UIP Clause Learning}
\label{sec:i-uip}
Although learning clauses with low LBD has been shown empirically to
be important in \sat solving than learning short clauses
\cite{DBLP:conf/ijcai/AudemardS09}, clause size is still
important. Smaller clauses consume less memory and help to decrease
the size of future learnt clauses. They are also semantically stronger
than longer clauses.

The all-UIP scheme will tend to produce small clauses since the
clauses contain at most one literal per decision level. However, the
all-UIP clause can have much higher LBD. Since, $\LBD$ is important
than size our approach is to use all-UIP learning when, and only when,
it succeeds in reducing the size of the clause \emph{without
  increasing its LBD}. The all-UIP scheme first computes the 1-UIP
clause when it reduces the deepest level to a single UIP literal. It
then proceeds to reduce the shallower levels (see all-UIP's for loop
in Figure~\ref{fig:cl_schemes}). So our approach will start with the
1-UIP clause and then try to apply all-UIP learning to reduce other
levels to single literals. As noted above, clause minimization is
orthogonal to our approach, so we also first apply standard clause
minimization \cite{DBLP:conf/sat/SorenssonB09} to the 1-UIP
clause. That is, our algorithm will start with the clause that most
\sat solvers learn from a conflict (a minimized 1-UIP clause). Our
approach is specified in Alg. ~\ref{alg:i-uip}.

\begin{algorithm}[!t]
\caption{\allUip}\label{alg:i-uip}
\begin{algorithmic}[1]
\Require $C_1$ is minimized 1-UIP clause
\Require $\tgap\geq 0$ is global parameter, $\ntries$ and $\nsuc$
         are globals used to configure $\tgap$
\Procedure{\allUip}{$C_1, \assertionTrail$, \type} 
    \State \textbf{if} ($|C_1| - \LBD(C_1) < \tgap$) \textbf{return} $C_1$\label{ln:control}
    \State $C_i \gets C_1$
    \State $\dlevels \gets\mbox{}$ decision levels in $C_i$ in descending order
           \Comment{These never change}
    \For{($i=1$; $i < |\dlevels|$; $i$++)} \Comment{skip the deepest level $\dlevels[0]$}
       \State $C_i\gets\mbox{}$\tryuiplevel($C_i$, $\dlevels[i]$, \type)
              \Comment{Try to reduce this level to UIP}
       \If{$\bigl|\{\ell\,|\,\ell \in C_i \land \dlevel{\ell} \geq \dlevels[i]\}\bigr|
                    + (|\dlevels| - (i+1)) \geq |C_1|$}\label{ln:earlyTerminate}
           \State \textbf{return} $C_1$. \Comment{can't generate smaller clause}
       \EndIf
    \EndFor
    \State \textbf{if} ($\type = \allUipPure$) $C_i \gets \mathbf{minimize}(C_i)$\label{ln:puremin}
    \State $\ntries$++\label{ln:try}
    \State \textbf{if} ($|C_i| < |C_1|$) $\nsuc$++, \textbf{return} $C_i$ 
           \label{ln:returnmin}\label{ln:suc}
    \State \textcolor{gray}{$\triangleright\mbox{}$ Use next line instead for} \allUipAct
    \State \textcolor{gray}{
           \textbf{if} ($|C_i| < |C_1| 
                        \land (\code{AvgVarAct}(C_i) > \code{AvgVarAct}(C_1))$)
                       $\nsuc$++, \textbf{return} $C_i$} \label{ln:activity}
    \State \textbf{return} $C_1$
%    \State \textbf{return} \textbf{if} $|C_i| < |C_1|$) 
%           \textbf{then}  $C_i$ \textbf{else}  $C_1$
\EndProcedure
\Statex
\Procedure{\tryuiplevel}{$C_i$, $i$, \type}\Comment{Do not add new decision levels}
    \State $\ctry = C_i$
    \State $L_i = \left\{\ell \left| \ell \in \ctry 
                         \land \dlevel{l} = i\right.\right\}$\label{ln:litsati1}
    \While{$|L_i| > 1$}
        \State $p \gets\mbox{}$ \textbf{remove} lit with the highest trail index from $L_i$ 
        \If{($\exists q \in \reason{\lnot p}.\, \dlevel{q} \not \in \mathit{decLvls}$)}\label{ln:newdlevel}
        \Comment{Would add new decision levels}
             \If{($\type = \allUipPure$)}
                 \State \textbf{return} $C_i$ \Comment{Abort, can't UIP this level}
             \ElsIf{($\type = \allUipMin$)}
                 \State \textbf{continue} \Comment{Don't try to resolve away $p$}
             \EndIf
        \Else\label{ln:nonewdlevel}
            \State $\ctry \gets \ctry \resolve \reason{\lnot p}$
            \State $L_i = L_i \cup \{\ell\,|\, 
                                     \ell\in\reason{\lnot p} \land \ell\neq \lnot p \land \dlevel{\ell} = i
                                    \}$\label{ln:litsati2}
        \EndIf
     \EndWhile
    \State \textbf{return} $\ctry$
\EndProcedure
\end{algorithmic}
\end{algorithm}

The Algorithm $\allUip$ attempts to reduce each level in the 1-UIP
clause to a single literal (a UIP). The 1-UIP clause has only one
literal at its deepest level, so that level, $\dlevels[0]$, can be
skipped. The subroutine \tryuiplevel is used to reduce each level. It
uses trail resoluions to achieve this, subject to the constraint that
no new decision levels can be introduced in the clause. In particular,
\tryuiplevel($C_i$, $i$) attempts to resolve away the literals at
decision level $i$ in the clause $C_i$, i.e., those in the set $L_i$
(line~\ref{ln:litsati1}), in order of decreasing trail index, until
only one literal at level $i$ remains. If the resolution step will not
introduce any new decision levels (line~\ref{ln:nonewdlevel}), it is
performed updating $\ctry$. In addition, all new literals at level $i$
are added to $L_i$. These are the literals at level $i$ in the reason
clause.

On the other hand, if the resolution step would introduce new decision
levels (line~\ref{ln:newdlevel}) then there are two options.  The
first option we call $\allUipPure$. With $\allUipPure$ we abort our
attempt to UIP this level and return the clause with level $i$
unchanged. In the second option, called $\allUipMin$, we continue
without performing the resolution, \textit{keeping} the current
literal $p$ in $\ctry$. $\allUipMin$ then continues to try to resolve
away the other literals in $L_i$ (note that $p$ is no longer in $L_i$)
until $L_i$ is reduced to a single literal. Hence, $\allUipMin$ can
leave multiple literals at level $i$---all of those with reasons
containing new levels along with one other.\footnote{Since the sole
  remaining literal $u\in L_i$ is at a lower trail index than all of
  the other literals, so there is no point in trying to resolve away
  $u$---either it will be the decision clause for level $i$ having no
  reason, or its reason will contain at least one other literal at
  level $i$.} Observe that the number of literals at level $i$ can not
be increased after processing it with \allUipPure. \allUipMin can,
however, potentially increase the number of literals at a level. In
resolving away a ltieral $l$ at level $i$ more literals might be
introduced into level $i$, and some of these might not be removable by
\allUipMin if their reasons contain new levels. However, both
\allUipPure and \allUipMin can increase the number of literals at
levels less that $i$ as new literals can be introduced into those
levels when the literals at level $i$ are resolved away. These added
literals at the lower levels might not be removable from the clause,
and thus both methods might yield a longer clause than the inputted
1-UIP clause.

After trying to UIP each level the clause $C_i$ is obtained. If we
were using $\allUipPure$ we can once again apply minimization
(line~\ref{ln:puremin}). Recursive clause minimization
\cite{DBLP:conf/sat/SorenssonB09} would be useless for the $\allUipMin$
clause as all but one literal of each level introduces a new level and
thus cannot be recursively removed.\footnote{Other more powerful
  minimization techniques could still be applied.}

Finally, the Algorithm returns the shorter of the initial 1-UIP clause
and the newly computed clause (line~\ref{ln:returnmin}). Because we
return the newly computed clause only when it is shorter an early
termination test can be used (line~\ref{ln:earlyTerminate}). After the
algorithm has finished processing levels $\dlevels[0]$--$\dlevels[i]$
the literals at those levels will not be further changed. Furthermore,
we know that the best that can be done is to reduce the remaining
$|\dlevels| - (i+1)$ levels down to a single literal each. These two
observations give a lower bound on the size of the resulting clause,
and if that lower bound is as large as the size of the initial 1-UIP
clause we can terminate and return the initial 1-UIP clause. As
observed above, \allUipPure cannot trigger this early termination
condition (it cannot increase the size of any processed level), but
\allUipMin can.

\paragraph{$\tgap$:}
$\allUip$ can produce significantly smaller clause. However, when it
does not yield a smaller clause, the cost of the additional resolution
steps can hurt the solver's performance. Since resolution cannot
reduce a clause's $\LBD$, The maximum size reduction obtainable from
$\allUip$ is the difference between the 1-UIP clause's size and its
$\LBD$: $\gap(C_1) = |C_1| - \LBD(C_1)$. When $\gap(C_1)$ is small,
applying $\allUip$ is unlikely to be cost effective. Our approach is
to dynamically set a threshold on $\gap(C_1)$, $\tgap$, such that when
$\gap(C_1) < \tgap$ we do not attempt to reduce the clause
(line~\ref{ln:control}). Initially, $\tgap = 0$, and we count the
number of times $\allUip$ is attempted ($\ntries$) and the number of
times it successfully yields a shorter clause ($\nsuc$)
(line~\ref{ln:try} and \ref{ln:suc}). On every restart if the success
rate is less than 80\%, we decrease $\tgap$ by one (not allowing it to
become negative), and if it is greater than 80\% we increase $\tgap$
by one.

\begin{example}
    Consider the trail $\assertionTrail = \ldots$, $\ell_1$, $a_2$, $b_2$,
    $c_2$,$d_2$, $\ldots$, $\ldots$, $e_5$, $f_5$, $g_5$,
    $h_6$, $i_6$, $j_6$, $k_6$, $\ldots$, $m_{10}$,
    $\ldots$ where the subscript indicates the decision level of each
    literal and the literals are in order of increasing trail
    index. Let the clauses
    $C_x$ denote the reason clause for literal $x_i$ with 
    \[\begin{array}{|l|l|l|}\hline
      C_{a} = \varnothing 
      & C_{b} =  (b_2, \lnot \ell_3, \lnot a_2)
      & C_{c} =  (c_2, \lnot a_2, \lnot b_2) \\\hline
      C_{d} =  (d_2, \lnot b_2, \lnot c_2) 
      & C_{\ell} =\varnothing
      & C_{e} =\varnothing \\\hline
      C_{f} =  (f_5, \lnot e_5,\lnot \ell_1)
      & C_{g} = (g_5, \lnot a_2, \lnot f_5)
      & C_{h} =\varnothing \\\hline
      C_{i} =  (i_6, \lnot e_5, \lnot h_6)
      & C_{j} = (j_6, \lnot f_5, \lnot i_6 )
      & C_{k} = (k_6, \lnot f_5, \lnot j_6) \\\hline
      \end{array}
  \]
  Suppose 1-UIP learning yields the clause
  $C_1 = (\neg{m_{10}}$, $\neg{k_6}$, $\neg{j_6}$, $\neg{i_6}$,
  $\neg{h_6}$, $\neg{g_5}$, $\neg{d_2}$, $\neg{c_2})$ where $\neg{m_{10}}$
  is the UIP from the conflicting level. $\allUip$ first tries to find
  the UIP for level 6 by resolving $C_1$ with $C_k$, $C_j$
  and then $C_i$ producing the clause
  $C^* = (\neg{m_{10}}$, $\neg{h_6}$, $\neg{g_5}$,
  $\neg{f_5}$, $\neg{e_5}$, $\neg{d_2}$, $\neg{c_2})$ where
  $\neg{h_6}$ is the UIP for level 6.

  $\allUip$ then attempts to find the UIP for level 5 by resolving $C^*$
  with $C_g$ and then $C_f$. However, resolving with $C_f$ would
  introduce $\ell_1$ and a new decision level into $C^*$. $\allUipPure$
  thus leaves level 5 unchanged . $\allUipMin$, on the other hand, skips
  the resolution with $C_f$  leaving $f_5$ in $C^*$. Besides $f_5$
  only one other literal at level 5 remains in the clause, $e_5$,
  so $\allUipMin$ does not do any further resolutions at this level. Hence,
  $\allUipPure$ yields $C^*$ unchanged, while $\allUipMin$ yeilds
  $C^*_{\mathit{min}} = (\neg{m_{10}}$, $\neg{h_6}$, $\neg{f_5}$,
  $\neg{e_5}$, $\neg{d_2}$, $\neg{c_2}, \lnot a_2)$.

  Finally, $\allUip$ processes level 2. Resolving away $d_2$ and then
  $c_2$ will lead to an attempt to resolve away $b_2$. But again this
  would introduce a new decision level with the literal $\ell_1$. So
  $\allUipPure$ will leave level 2 unchanged and $\allUipMin$ will leave
  $b_2$ unresolved. The final clauses produced by $\allUipPure$ would be
  $(\neg{m_{10}}$, $\neg{h_6}$, $\neg{f_5}$, $\neg{e_5}$, $\neg{d_2}$,
  $\neg{c_2}$, $\neg{a_2})$, a reduction of 1 over the 1-UIP clause,
  and by $\allUipMin$ would be $(\neg{m_{10}}$, $\neg{h_6}$,
  $\neg{f_5}$, $\neg{e_5}$, $\neg{b_2}$, $\neg{a_2})$, a reduction of
  2 over the 1-UIP clause.
\whitebox
\end{example}

\subsection{Variants of $\allUip$}
We also developed and experimented with a few variants of the
$\allUip$ algorithm which describe below.

\subsubsection{$\allUipAct$: Clauses with Active
  Variables.} \label{sec:active} $\allUip$ learning might introduce
literals with low variable activity into $C_i$. Low activity variables
are variables that have had low recent participation in clause
learning. Hence, clauses with variables of low activity might not be
as currently useful to the solver. Our variant $\allUipAct$, shown as
the optional line~\ref{ln:activity} in Algorithm~\ref{alg:i-uip},
computes the average variable activity of the newly produced all-UIP
clause $C_i$ and the original 1-UIP clause $C_1$. The new clause $C_i$
will be returned only if it is both smaller and has higher average
variable activity than the original 1-UIP clause. There are, of
course, generalizations of this approach where one has a weighted
tradeoff between these factors that allows preferring the new clause
when it has large gains in one metric even though it has small losses
in the other. We did not, however, experiment with such
generalization.

\subsubsection{Adjust Variable Activity.} \label{sec: varajust} An
alternative to filtering clauses with low average variable
($\allUipAct$) is to alter the way variable activities are updated to
account for our new clause learning method. The popular branching
heuristics VSIDS \cite{DBLP:conf/dac/MoskewiczMZZM01} and LBR
\cite{DBLP:conf/sat/LiangGPC16} bump the variable activity for all
literals appearing in the learnt clause $C_L$ and all literals
resolved away during the conflict analysis that yielded $C_L$ from the
initially detected conflict $C_I$ (all literals on the conflict side).

We did not apply this approach to the $\allUip$ clause, as we did not
want to bump the activity of the literals at above the deepest
decision level that $\allUip$ resolves away. Intuitively, these
literals did not directly contribute to generating the
conflict. Instead, we tried two modifications to the variable activity
bumping schemes.

Let $C_1$ be the 1-UIP learnt clause and $C_i$ be the $\allUip$ learnt
clause. First, we kept all of the variable activity bumps normally
done by 1-UIP learning.\footnote{So extra techniques used by the
  underlying solver, like reason side rate and locality
  \cite{DBLP:conf/sat/LiangGPC16}, were kept intact.}  Then, when the
$\allUip$ scheme was successful, i.e., $C_i$ was to be used as the new
learnt clause, we perform further updates to the variable
activities. In the $\allUipIn$ approach all variables variables
appearing in $C_i$ that are not in $C_1$ have their activities
bumped. Intuitively, since the clause $C_i$ is being added to the
clause database we want to increase the activity of all of its
variables. On the other hand, in the $\allUipEx$ approach in addition
to bumping the activity of the new variables in $C_i$ we also remove
the activity bumps of those variables in $C_1$ that are no longer in
$C_i$.

In sum, the two modified variable activity update schemes we
experimented with were
\[
    \begin{array}{l}
    \textbf{\allUipIn} \equiv \forall l\in C_i - C_1.\,\mathit{bumpActivity}l) \\
    \textbf{\allUipEx} \equiv \forall l\in C_i - C_1.\,\mathit{bumpActivity}(l) \land 
    \forall l\in C_1 - C_i.\,\mathit{unbumpActivity(l)}
  \end{array}
\]

\subsubsection{Chronological Backtracking.}
We tested our new clause learning schemes on solvers that utilized
Chronological Backtracking
\cite{DBLP:conf/sat/NadelR18,DBLP:conf/sat/MohleB19}. When
chronological backtracking is used the literals on the trail might no
longer be sorted by decision level. So resolving literals in the
conflict by highest trail index first no longer works. However, we can
define a new ordering on the literals to replace the trail index
ordering. Let $l_1$ and $l_2$ be two literals on the trail
$\trail$. We say that $l_1 >_{\mathit{chron}} l_2$ if
$\dlevel{l_1} > \dlevel{l_2} \lor (\dlevel{l_1} = \dlevel{l_2} \land
\trailIdx{l_1} > \trailIdx{l_2})$. That is, literals with higher
decision level come first, and if that is equal then the literal with
higher trail index comes first.

Exploiting the analysis of \cite{DBLP:conf/sat/MohleB19} it can be
observed that all clause learning schemes continue to work as long
literals are resolved away from the initial conflict in decreasing
$>_{\mathit{chron}}$ order. In our implementation we used a heap
(priority queue) to achieve this ordering of the literal resolutions
in order to add our new schemes to those solvers using chronological
backtracking.

\section{Implementation and Experiments}
\subsubsection{Clause Reduction with $\allUip$.}
We implemented our new clause learning schemes in the
\text{\MapleBase} \cite{}, the winner of SAT Race 2015 application
track. We then evaluated our schemes on the full set of benchmarks
from SAT RACE 2019 main track which contains 400 instances. We ran our
experiments on \fb{insert GHz of CPU}, allowing \fb{insert time bound}
seconds per instance and a maximum of \fb{insert RAM bound}.

\fb{Delete figure 2 and use only figure 8 with all of the
  solvers. First briefly comment on the performance and make a brief
  observation about clause sizes.

  Then have a new subsubsection that looks at clause sizes wrt. to the
  two schemes \allUipPure and \allUipMin. Shrink fig 3 and 4 and put
  them side by side (using a tabular environment as I did in figure 1).

  Remove fig 5 and put those numbers in the text instead of in a table
  as part of your discussion.

  Then have a new subsubsection on DRAT trim and get rid of Fig 7.

  Then a subsubsection on \allUip in modern solvers. Some of our
  observations have no solid foundation, so check them to make sure
  that you are just not stating your intuitions without reasonable
  evidence. We need to to fit into 15 pages total along with a short
  1/3 page for conclusions.

  Finally add all citations.

  If we add the more complex \allUipMin we also will need more space
  to describe that, so prune some of your conclusions especially those
  that are simply conjectures.
}

Beside solved instances count and PAR-2 score, we additionally measure
the average clause length and clause reduction ratio (both cumulative
and non-cumulative)\footnote{The cumulative reduction ratio is
  obtained through learning all clauses with the target learning
  scheme; Therefore, the reduction is cumulative. The non-cumulative
  reduction ratio is obtained by running the target scheme for
  measurement only (the minimized 1-UIP clause is learned); Therefore,
  the reduction is not cumulative. }  for each instances. For
$\MapleIUIPPure$ and $\MapleIUIMin$, we also captures the $\allUip$
learning attempted rate and success rate.

\begin{figure} 
\begin{center}
\begin{tabular}{ | m{3.3cm} | m{2cm}| m{2cm} | m{2cm} | m{2.7cm} | } 
\hline
Solver & \# solved & PAR-2 & Clause Size & Cl Reduction\% \\ 
\hline
$\MapleBase$ & 221 (132, 89) & 5018.89 & 62.6 & 36.53\% \\ 
\hline
$\MapleIUIPPure$ & \textbf{228} (135, 93) & \nf{4867.37} & 49.88 & 41.6\%  (42.72\%)\\ 
\hline
$\MapleIUIMin$ & 226 (135, 91) & 4890.67 & \textbf{45.2} & \textbf{47.8\% (51.19\%)}\\ 
\hline
\end{tabular}
\end{center}
\caption{Benchmark results of $\MapleBase$ , $\MapleIUIPPure$  and $\MapleIUIMin$ on SAT2019 race main track.
CL Reduction\% is the clause size reduction ratio comparing to non-minimized 1-UIP clauses, and the values in the brackets are the non-cumulative reduction ratio.}
\label{fig:t1}
\end{figure}

Fig.~\ref{fig:t1} shows that $\allUipPure$ ($\allUipPure$) solved seven
(five) more instances than the baseline solver with lower PAR-2
scores. $\allUipPure$ has marginally lower PAR-2 score than
$\allUipPure$. Both $\allUipPure$ and $\allUipMin$ produce clause with
significantly smaller size than 1-UIP by 20.4\% and 27.7\%,
respectively. Fig.~\ref{fig:len_pdf} shows the probability density
distribution (PDF) of the relative clause size of both i-uip learning
schemes ($\allUipPure$ in red and $\allUipMin$ in green) for each
instance. $\allUipPure$ ($\allUipMin$) produces shorter clauses for 77.7\%
(88.5\%) of instances, and average relative reduction from 1-UIP is
16.7\% (18.6\%). Fig.~\ref{fig:len_compare} compares the absolute
clause size of $\allUipMin$, $\allUipPure$ and 1-UIP, and it shows that
both i-UIP learning schemes in general produce the smaller clauses,
and the size reduction is more significant for instances with large
average 1-UIP clause size. Moreover, $\allUipMin$ clauses (indicated in
green) is consistently smaller than $\allUipPure$ clauses.

2\begin{figure}
    \centering
    \includegraphics[width=0.8\textwidth]{figures/clause_reduction_PDF}
    \caption{ Relative clause size reduction distribution. X axis
      indicates the relative size of difference between i-uip and
      1-uip clauses (calculated as
      $\dfrac{|C_1|-|C_i|}{|C_1|}$ ) for
      each instance, and Y axis shows the probability density.}
     \label{fig:len_pdf}
\end{figure}
\begin{figure} \label{fig:len_compare}
    \centering
    \includegraphics[width=0.8\textwidth]{figures/i-uip-sizes-compare.pdf}
    \caption{Average clause size comparison plot. Each point in the
      plot represents an instance. X and Y axis shows the clause
      length from $\allUip$ and 1-UIP, respectively. Each green (red)
      dot represents an compared instance between $\MapleBase$ and
      $\MapleIUIMin$ ($\allUipPure$). }
    \label{fig:len_compare}
\end{figure}
\nf{ Do we need this?  We additionally looked at the 14 instances
  solved by $\allUipMin$ but not by 1-UIP. $\allUipMin$ produces smaller
  clauses for all of them with average relative reduction of 22\% and
  maximum 77\% (30 vs 135). Seven out of 14 instances has size
  relative reduction over 30\%. For the 9 instances solved by 1-UIP
  but not by $\allUipMin$, $\allUipMin$ only produce smaller clause for
  three instances and with average relative reduction of 3.3\%.}

$\allUipMin$ outperformed $\allUipPure$ in clause size.  This results agrees with our observation in Fig.~\ref{fig:t2}: $\allUipMin$ attempted $\allUip$ learning more frequently, and it is more likely to succeed.

\begin{figure} 
\begin{center} 
\begin{tabular}{ | m{3.5cm} | m{5cm}| m{3.5cm} | } 
\hline
Solver & $\allUip$ attempt rate & $\allUip$ success rate  \\ 
\hline
$\MapleIUIPPure$ & 16.1\% & 43.4\% \\ 
\hline
$\MapleIUIMin$ & 28.8\% & 59.3\% \\ 
\hline
\end{tabular}
\end{center}
\caption{Compare $\allUipPure$ and $\allUipMin$ i-uip attempt rate and success rate. $\allUipMin$ scheme attempted $\allUip$ more frequently, and it is more likely to successfully produce smaller $C_i$ clause .}
\label{fig:t2}
\end{figure}

A solver produce smaller clauses can construct smaller proofs. For
UNSAT instances, we measure their DRUP\cite{} proof checking time as
well as the size of the optimized DRUP proof. We used DART-trim
\cite{} with 5000 timeout to check and optimize DRUP proofs.

Fig.~\ref{fig:t3} shows that the optimized proof construct by
$\allUipMin$ and $\allUipPure$ are significantly smaller than 1-UIP
proofs. The relative proof size reduction roughly correlates to the
average clause size reduction. Fig.~\ref{fig:proof_compare} shows the
absolute proof size comparison results.

\begin{figure} 
\begin{center} 
\begin{tabular}{ | m{3.5cm} | m{5cm}| m{3.5cm} | } 
\hline
Solver & optimized proof size (MB) & relative reduction size  \\ 
\hline
$\MapleBase$ & 613.9 & 0  \\ 
\hline
$\MapleIUIPPure$ & 487.2 & 6.90\% \\ 
\hline
$\MapleIUIMin$ & 413.2 & 17.18\% \\ 
\hline
\end{tabular}
\end{center}
\caption{Optimized UNSAT proof comparison for 1-UIP $\allUipPure$ and $\allUipMin$. Optimized proof size measures the average absolute proof size in MB, and relative reduction size measures the average relative reduction for all UNSAT instances.}
\label{fig:t3}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=0.8\textwidth]{figures/proof_size_compare.pdf}
    \caption{Average optimized proof size between 1-uip and $\allUipMin$.}
    \label{fig:proof_compare}
\end{figure}

\begin{figure} 
\begin{center}
\begin{tabular}{ | m{3.5cm} | m{4cm}| m{2cm} | m{2.75cm} |  } 
\hline
Solver & \# solved (SAT, UNSAT) & PAR-2 & Avg clause Size \\ 
\hline
1-UIP & 221 (132, 89)  & 5018.89 & 62.6  \\ 
\hline
\nf{$\allUipPure$} &\textbf{228} (135, \nf{93}) & 4867.37 & 49.88 \\
\hline
$\allUipMin$ & 226 (135, 91) & 4890.67 & 45.2 \\ 
\hline
$\allUipAct$ & 226 (135, 91)  & \textbf{4866.94} & 47.7 \\
\hline
$\allUipIn$ & 225 (\textbf{138}, 87) & 4958.49 & 52..12 \\
\hline
$\allUipEx$ & 223 (134, 89) & 5015.23 & \textbf{43.2} \\
\hline
\end{tabular}
\end{center}
\caption{Benchmark results of $\MapleBase$ with 1-UIP, $\allUipPure$, $\allUipMin$, $\allUipAct$,
$\allUipIn$, and $\allUipEx$ on SAT2019 race main track.}
\label{fig:t4}
\end{figure}

\subsection{$\allUip$ as a Practical Learning Scheme}
To evaluate $\allUip$'s effectiveness as a clause learning scheme, we
re-implement $\allUipMin$ on $\MapleBase$ with the extensions mentioned
in section~\ref{sec:i-uip}. We evaluated 1-UIP learning and five
$\allUip$ configurations ( $\allUipMin$, $\allUipPure$, $\allUipAct$,
$\allUipIn$, and $\allUipEx$) on the SAT Race 2019 main track
benchmark and reported solved instances, PAR-2 score and average
clause size.

Fig.~\ref{fig:t4} summarizes the result of the experiment. Learning
scheme $\allUipPure$ solved the most overall instances (228) and the
most UNSAT instances (93). $\allUipAct$ had the lowest PAR-2
score. $\allUipIn$ solved the most SAT instances (138). $\allUipEx$
produced the shortest average clause size, but solved the second least
instances, one more instance than the baseline 1-UIP learning. All
configurations of $\allUip$ outperformed the baseline 1-UIP scheme in
solved instances, PAR-2 score and average clause size.

\subsection{$\allUip$ on Modern SAT solvers}
To validate $\allUip$ as a generalizable learning scheme on modern SAT
solvers, we re-implement $\allUip$ on the winners of 2017, 2018 and 2019
SAT Race\cite{} and $\expSAT$\cite{} ($\expSATShort$). $\expSATShort$
is a top ten solver from 2019 SAT race which uses random walk
simulation to help branching. We chose $\expSATShort$ because 1) it is
a top solver in the 2019 SAT Race without using chronological
backtracking; 2) the combination of random walk simulation and
variable activity branching heuristic allows our learning schemes to
partially sidestep the problem of variable activity. For each solver,
we compare the base 1-UIP learning scheme against $\allUipPure$,
$\allUipMin$ and the top two $\allUip$ variants, $\allUipAct$ and
$\allUipIn$, on the SAT Race 2019 main track benchmark. We report
solved instances, PAR-2 score and the average clause size.

\begin{figure} 
\begin{center}
\begin{tabular}{ | m{3.7cm} | m{4cm}| m{2cm} | m{2.75cm} |  } 
\hline
Solver & \# solved (SAT, UNSAT) & PAR-2 & Avg clause Size \\ 
\hline
SAT 2017 Winner & & & \\
$\MapleSeven$ & 232 (135, 97) & 4755.96 & 61.9  \\ 
\hline
$\MapleSeven$-i-pure & \textbf{244 (146, 98)} +12 & \textbf{4504.18} & 43.76 \\
\hline
$\MapleSeven$-i-min & 240 (144, 96) +8 & 4601.25 & \textbf{36.97} \\ 
\hline
$\MapleSeven$-i-Act & 237 (140, 97) +5 & 4678.434 & 43.62 \\ 
\hline
$\MapleSeven$-i-inclusive & 234 (137, 97) +2 & 4718.03 & 37.96 \\
\hline
\hline
SAT 2018 Winner & & & \\
$\MapleEightShort$ & 236 (138, 98) & 4671.81 & 61.69 \\
\hline
$\MapleEightShort$-i-pure & \textbf{241} (\textbf{142, 99}) +5 & \textbf{4598.18} & 44.19 \\
\hline
$\MapleEightShort$-i-min & 236 (141, 95) +0 & 4683.92 & 38.05 \\ 
\hline
$\MapleEightShort$-i-Act & 240 (141, \textbf{99}) +4 & 4626.99 & 41.16 \\
\hline
$\MapleEightShort$-i-inclusive & 240 (\textbf{142}, 98) +4 & 4602.13 & \textbf{37.52} \\
\hline
\hline
SAT 2019 Winner & & & \\
$\MapleNineShort$ & 238 (140, \textbf{98}) & 4531.24 & 60.91 \\
\hline
$\MapleNineShort$-i-pure & 238 (140, \textbf{98}) +0 & 4519.08 &  43.32\\
\hline
$\MapleNineShort$-i-min & 244 (\textbf{148}, 96) +6 & \textbf{4419.84} & \textbf{36.88} \\
\hline
$\MapleNineShort$-i-Act & 243 (146, 97) + 5 & 4476.73 & 40.65 \\
\hline
$\MapleNineShort$-i-inclusive & \textbf{243} (\textbf{148}, 95) +5 & 4455.76 & 37.02 \\
\hline
\hline
SAT 2019 Competitor & & &\\
$\expSATShort$ & 237 (137, 100)  & 4628.96 & 63.19 \\
\hline
$\expSATShort$-i-pure & 235 (136, 99) -2  & 4668.96 & 48.26 \\
\hline
$\expSATShort$-i-min & 241 (143, 98) +4 & 4524.28 & 46.29 \\ 
\hline
$\expSATShort$-i-Act & 244 (143, \textbf{101}) +7 & \textbf{4460.92} & 47.25 \\
\hline
$\expSATShort$-i-inclusive & \textbf{245} (\textbf{146}, 99) +8 & 4475.76 & \textbf{45.33} \\
\hline
\end{tabular}
\end{center}
\caption{Benchmark results of 1-UIP, $\allUipPure$. $\allUipMin$, $\allUipAct$ and $\allUipIn$ on SAT2019 race main track.}
\label{fig:t5}
\end{figure}

Table~\ref{fig:t5} shows the benchmark result of $\allUip$
configurations on different solvers. All four configurations of
$\allUip$ outperformed 1-UIP learning for the SAT 2017 race winner,
$\MapleSeven$. More specifically, $\allUipPure$, $\allUipMin$,
$\allUipAct$ and $\allUipIn$ solved 12, 8, 5 and 2 more instances,
respectively, whiling producing smaller clauses. $\allUipPure$ solved
more UNSAT and SAT instances while other configurations improved on
solving SAT instances.  The improvement of $\allUip$ is more significant
on $\MapleSeven$ than on $\MapleBase$ for both solved instances and
clause size reduction. This may suggests that $\allUip$ and the recent
learnt clause minimization approach~\cite{} synergies well because
i-UIP clauses are shorter with more common literals which allows
vivification~\cite{} to prune literals more aggressively through unit
propagation.

We observed significant improvement of $\allUip$ for the SAT 2018
winner, $\MapleEightShort$. Three out of four $\allUip$ configurations
outperformed 1-UIP by 5, 4 an 4 instances,
respectively. $\MapleEightShort$ improved from $\MapleSeven$ by using
chronological backtracking (CB) for long distance
backtracks. Therefore, we used $\allUip$-CB extension for all $\allUip$
configurations. The results indicates that the improvement of $\allUip$
is slightly shadowed by the adoption of CB. More specifically, we
believe CB prevents decision levels from being compressed through the
process of backtracking and literal assertion. The shorter i-UIP
clauses can bring related literals closer, which in turns compresses
the assertion stack and the decision levels. However, since CB
discourages long distance backtracking, the effect of learning shorter
clauses is shadowed until a full restart.  We believe we have observed
an interesting interaction effect that shows the limitations of both
CB and $\allUip$ learning for future research.

$\allUip$ learning schemes showed significant improvement for
$\MapleNineShort$. Three out of four configurations improved solved
instances by 6, 5 and 5 instances, respectively. $\MapleNineShort$
uses duplicate learning (DL) to prioritize clauses that are learned
multiple times. $\allUip$ and DL didn't synergies well possibly because
i-UIP clauses are less likely to be duplicated. We observed that
$\allUipMin$ ,in average, added 12\% less duplicated clauses into the
core clause database than 1-UIP. One possible explanation is that
$\allUip$ learning can reduce an 1-UIP clause to different i-UIP clauses
for different assertion trails. Therefore, the solver is less likely
to learn the same i-UIP clauses.

We observed significant improvement of $\allUip$ for $\expSATShort$,
three out of four configurations of $\allUip$ significantly outperformed
1-UIP learning by 4, 7 and 8 more instances, respectively, while
producing smaller clauses. We expect all three configurations of
$\allUip$ to solve the similar amount of instances with close PAR-2
scores because the additional random walk exploration allows the
learning schemes to partially sidestep the activity problem; hence,
the learning adjustment for variable activities should have less
impact on the solver's performance. However, we instead, observed that
both $\allUipAct$ and $\allUipIn$ outperformed the default i-UIP
learning scheme. One possible explanation is that $\allUipAct$ and
$\allUipIn$ schemes could easily overcompensate variable activities
for i-UIP clauses, and $\expSATShort$'s random walk exploration could
use future search information to mitigate the negative effects of our
overcompensation.

\bibliography{sat}{}
\bibliographystyle{splncs04}
\end{document}